Higher-Order unification for free!

### 2024-04-10 paper/main.tex
typo
update ref

\begin{lemma}
Properties \ref{prop:correct} and
\ref{prop:complete} hold for the implementation of \Uo above
\end{lemma}
\begin{proof}[Proof sketch]
 In this setting \Ee is as strong as
\Eo on ground terms. What we have to show is that whenever two different \Fo
terms can be made equal by a substitution $\rho$ (plus the \ref{clause:beta1}
and \ref{clause:beta2} if needed) we can find this $\rho$ by finding
a $\sigma$ via \Ue{} on the corresponding \Ho terms and by decompiling it.
If we look at the \Fo{} terms, the are two interesting cases:
\begin{itemize}
\item \elpiIn{fuva X ~\Uo~s}. In this case after \elpiIn{comp} we have
  $Y \Ue t$ that succeeds with $\sigma = \{ Y \mapsto t\}$ and
  $\sigma$ is decompiled to $\rho = \{ Y \mapsto s\}$.
\item \elpiIn{fapp[fuva X|L] ~\Uo~s}. In this case
 we have $Y_{\vec{x}} \Ue t$ that succeeds with
 $\sigma = \{ \vec{y} \vdash Y \mapsto t[\vec{x}/\vec{y}]\}$ that in turn
 is decompiled to $\rho = \{ Y \mapsto \lambda \vec{y}.s[\vec{x}/\vec{y}]\}$.
 Thanks to \ref{clause:beta1}
 $(\lambda \vec{y}.s[\vec{x}/\vec{y}])~\vec{x} \Eo s$.
\end{itemize}
Since the mapping is a bijection occur check in \Ho{} corresponds to occur
check in \Fo{}.
\end{proof}

@@

\subsection{Relaxing lemma \nameref{lemma:keep-fidelity}}

@@

A second strategy to deal with problem that are in \notllambda is to make some
approximation. This is the case for example of the unification algorithm of Coq
used in its type class solver~\cite{sozeau08}. The approximation consists in
forcing a choice (among the others) when the unification problem is in
\notllambda. For instance, in $X \appsep a \appsep b = Y \appsep b$, the last
argument of the two terms is the same, therefore $Y$ is assigned to $X a$. Note
that this is of course an approximation, since $\sigma = \{X = \lambda x.Y, Y =
\_\}$ is another valid substitution for the original problem. This approximation
can be easily introduced in our unification procedure, by adding new custom 
\linkbeta progress rules.

>>>>

\begin{lemma}
Properties \ref{prop:correct-ml} and
\ref{prop:complete-ml} hold for the implementation of \Uo above
\end{lemma}
\begin{proof}[Proof sketch]
 In this setting \Ee is as strong as
\Eo on ground terms. What we have to show is that whenever two different \Fo{}
terms can be made equal by a substitution $\rho$ (plus the \ref{clause:beta1}
and \ref{clause:beta2} if needed) we can find this $\rho$ by finding
a $\sigma$ via \Ue{} on the corresponding \Ho terms and by decompiling it.
If we look at the \Fo{} terms, the are two interesting cases:
\begin{itemize}
\item \elpiIn{fuva X ~\Uo~s}. In this case after \elpiIn{comp} we have
  $Y \Ue t$ that succeeds with $\sigma = \{ Y \mapsto t\}$ and
  $\sigma$ is decompiled to $\rho = \{ Y \mapsto s\}$.
\item \elpiIn{fapp[fuva X|L] ~\Uo~s}. In this case
 we have $Y_{\vec{x}} \Ue t$ that succeeds with
 $\sigma = \{ \vec{y} \vdash Y \mapsto t[\vec{x}/\vec{y}]\}$ that in turn
 is decompiled to $\rho = \{ Y \mapsto \lambda \vec{y}.s[\vec{x}/\vec{y}]\}$.
 Thanks to \ref{clause:beta1}
 $(\lambda \vec{y}.s[\vec{x}/\vec{y}])~\vec{x} \Eo s$.
\end{itemize}
Since the mapping is a bijection occur check in \Ho{} corresponds to occur
check in \Fo{}.
\end{proof}

@@

\subsection{Relaxing \fullref{def:progBetaLH}}

@@

A second strategy to deal with problem that are in \notllambda is to make some
approximation. This is the case for example of the unification algorithm of Coq
used in its type class solver~\cite{sozeau08}. The approximation consists in
forcing a choice (among the others) when the unification problem is in
\notllambda. For instance, in $X \appsep a \appsep b = Y \appsep b$, the last
argument of the two terms is the same, therefore $Y$ is assigned to $X a$. Note
that this is of course an approximation, since $\sigma = \{X \mapsto \lambda x.Y, Y \mapsto
\_\}$ is another valid substitution for the original problem. This approximation
can be easily introduced in our unification procedure, by adding new custom 
\linkbeta progress rules.
